%*******************************************************
% Acknowledgments
%*******************************************************
\pdfbookmark[1]{Quotes}{quotes}

\thispagestyle{empty}

\begin{flushright}{\slshape
	\textit{By striving to do the impossible, man has always achieved what is possible. Those who have cautiously done no more than they believed possible have never taken a single step forward} \\
	Mikhail A. Bakunin\\*[0.5cm]
	\textit{It is only those who do nothing who makes no mistake} \\
	Pyotr Kropotkin\\*[0.5cm]
	\textit{A man provided with paper, pencil, and rubber, and subject to strict discipline, is in effect a universal machine} \\
	Alan Turing\\*[0.5cm]
	\textit{Don't try to solve serious matters in the middle of the night} \\
	Philip K. Dick\\	
	}
\end{flushright}



\newpage

\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax

\chapter{Acknowledgements}

{\small
First of all I would like to thank my supervisors: Professor \textbf{Alessio Mengoni} and Doctor \textbf{Anna Benedetti} for their constant presence and contagious enthusiasm.

\noindent I also would like to thank the COMBO (Florence computational biology group) in the persons of \textbf{Marco Galardini}, \textbf{Marco Fondi} and \textbf{Emanuele Bosi} for their advices and all discussions made in front of an old whiteboard.

\noindent Special thanks to Professor \textbf{Marco Bazzicalupo} and Professor \textbf{Alberto Ugolini} for gave me the opportunity to join in their research activities.

\noindent Many thanks to \textbf{Francesco Vitali} for sharing with me the experience of a PhD in the Department of Biology of the Florence University.

\noindent Thanks to all of those who have gravitated around the Specola and Sesto Fiorentino during these years, and especially \textbf{Alice Checcucci}, \textbf{Isabel Maida}, \textbf{Elena Perrin}, \textbf{Angela Frascella}, \textbf{Giulia Spini} and \textbf{Amaranta Focardi} for all the good laughs.

\noindent Many thanks to my family which has always supported me in all these long years of neverending studies: I wouldn't have made it without your presence.

\noindent Last but not least, thanks to \textbf{Adele Giannetti} for standing by me every time I was not-so-enthusiastic about what was going on.
}

\endgroup



