%*******************************************************
% Publications
%*******************************************************
\pdfbookmark[2]{Publications}{publications}

\chapter{Publications list}

Most of the data presented in this thesis has been published in the following \textit{peer-reviewed} papers:
\begin{itemize}

\item \textbf{Bacci, G.}, Bazzicalupo, M., Benedetti, A., \& Mengoni, A. (2014). StreamingTrim 1.0: a Java software for dynamic trimming of 16S rRNA sequence data from metagenetic studies. \textit{Molecular ecology resources}, 14(2), 426-434.

\item Mengoni, A., Focardi, A., \textbf{Bacci, G.}, \& Ugolini, A. (2013). High genetic diversity and variability of bacterial communities associated with the sandhopper \textit{Talitrus saltator} (Montagu)(Crustacea, Amphipoda). \textit{Estuarine, Coastal and Shelf Science}, 131, 75-82.

\item \textbf{Bacci, G.}, Pagoto, E., Passaponti, M., Vannocci, P., Ugolini, A., \& Mengoni, A. (2014). Composition of supralittoral sediments bacterial communities in a Mediterranean island. \textit{Annals of Microbiology}, 1-13.

\item Borruso, L., \textbf{Bacci, G.}, Mengoni, A., De Philippis, R., \& Brusetti, L. (2014). Rhizosphere effect and salinity competing to shape microbial communities in \textit{Phragmites australis} (Cav.) Trin. ex-Steud. \textit{FEMS microbiology letters}, 359(2), 193-200.

\item Bevivino, A., Paganin, P., \textbf{Bacci, G.}, Florio, A., Pellicer, M. S., Papaleo, M. C., ... \& Dalmastri, C. (2014). Soil Bacterial Community Response to Differences in Agricultural Management along with Seasonal Changes in a Mediterranean Region. \textit{PloS one}, 9(8), e105515.

\item Canfora, L., \textbf{Bacci, G.}, Pinzari, F., Papa, G. L., Dazzi, C., \& Benedetti, A. (2014). Salinity and bacterial diversity: to what extent does the concentration of salt affect the bacterial community in a saline soil?. \textit{PloS one}, 9(9), e106662.

\item \textbf{Bacci, G.}, Bani, A., Bazzicalupo, M., Ceccherini, M.T., Galardini, M., Nannipieri, P., Pietramellara, G., \& Mengoni, A. (2015). Evaluation of the performances of Ribosomal Database Project (RDP) Classifier for taxonomic assignment of 16S rRNA metabarcoding sequences generated from Illumina-Solexa NGS. \textit{Journal of Genomics}, 3, 36-39.

\item \textbf{Bacci, G.}, Ceccherini, M.T., Bani, A., Bazzicalupo, M., Castaldini, M., Galardini, M., Giovannetti, L., Mocali, S., Pastorelli, R., Pantani, O.L., Arfaioli, P., Pietramellara, G., Viti, C., Nannipieri, P., \& Mengoni, A. Exploring the dynamics of bacterial community composition in soil: the pan-bacteriome approach. Manuscripts in press in: \textit{Antonie van Leeuwenhoek}.

\end{itemize}

% Book chapter
\noindent \textit{Book chapter}
\begin{itemize}

\item \textbf{Bacci, G.} (2015). Raw Sequence Data and Quality Control. In \textit{Bacterial Pangenomics} (pp. 137-149). Springer New York.

\end{itemize}


% Submitted for publication
\noindent \textit{Submitted for publication}

\begin{itemize}

\item \textbf{Bacci, G.}, Abdelrhman, K.F.A., Marras, B., Nistri, A., Schintu, M., Ugolini, A., \& Mengoni, A. The gut microbiota of talitrid amphipods provides insight into the ecology of supralittoral sediments detritivors. Manuscripts submitted to: \textit{FEMS Microbiology Ecology}.

\item Paganin, P., Fiscarelli, E.V., Tuccio, V., Chiancianesi, M., \textbf{Bacci, G.}, Morelli, P., Dolce, D., Dalmastri, C., De Alessandri, A., Lucidi, V., Taccetti, G., Mengoni, A., \& Bevivino, A. Changes in Cystic Fibrosis Airway Microbial Community Associated with a Severe Decline in Lung Function. Manuscripts submitted to: \textit{PloS one}.

\item Galardini, M., Spini, G., Rossi, M., Roncaglia, B., Bani, A., \textbf{Bacci, G.}, Pini, F., Brilli, M., Biondi, E.G., Bazzicalupo, M., \& Mengoni, A. Evolution of intra-specific regulatory networks in \textit{Sinorhizobium meliloti}. Manuscripts submitted to: \textit{PLOS Computational Biology}

\end{itemize}


% Other articles in preparation related to this thesis
\textit{In preparation}

\begin{itemize}

%\item Di Carli M., De Rossi P., Paganin P., Del Fiore A., Lecce F., Capodicasa C., Bianco L., Perrotta G., Mengoni A., \textbf{Bacci, G.}, Benvenuto E., Vitali F., Daroda L., Dalmastri C., Donini M., Bevivino A. Bacterial community composition and proteome content of fresh-cut lettuce as affected by packaging.

\item \textbf{Bacci, G.}, Paganin, P., Lopez, L., Vanni, C., Dalmastri, C., Cantale, C., Daddiego, L., Perrotta, G., Taccetti, G., De Alessandri, A., Fiscarelli, E.E., Lucidi, V., Bevivino, A., \& Mengoni, A. Taxonomic signatures of CF airway microbiota distinguish between stable and severe declining patients. Manuscripts to be  submitted to: \textit{Proceedings of the National Academy of Sciences}


\end{itemize}


% Conference papers
\textit{Conference papers}

\begin{itemize}

\item \textbf{Bacci, G.} (2012, June). \textit{Elementi di Bioinformatica}. Paper presented at Scuola di Biodiversità e Bioindicazione della Società Italiana Scienza del Suolo. Rome, Italy.

\item \textbf{Bacci, G.}, Bazzicalupo, M., Benedetti, A., \& Mengoni, A. (2013, October). \textit{StreamingTrim 1.0: a Java software for dynamic trimming of 16S rRNA sequence data from metagenetic studies}. Paper presented at Bioinformatiha 2. Florence, Italy.

\item \textbf{Bacci, G.}, Bazzicalupo, M., Benedetti, A., \& Mengoni, A. (2013, December). \textit{StreamingTrim 1.0: a Java software for dynamic trimming of 16S rRNA sequence data from metagenetic studies}. Poster presented at 2\textsuperscript{nd} T\"{u}nen symposium on soil Metagenomics: mining and learning from metagenomes. Braunschweig, Germany.

\item Canfora, L., \textbf{Bacci, G.}, Pinzari, F., Papa, G. L., Dazzi, C., \& Benedetti, A. (2013, December). \textit{Metagenomic approach applied to the microbial community of a saline soil with high spatial variability}. Poster presented at 2\textsuperscript{nd} T\"{u}nen symposium on soil Metagenomics: mining and learning from metagenomes. Braunschweig, Germany.

\item \textbf{Bacci, G.} (2014, May). \textit{Inspecting the dynamics of bacterial communities. The pan-bacteriome approach}. Paper presented at Cortona Procarioti. Cortona, Italy.

\item \textbf{Bacci, G.}, Bazzicalupo, M., Benedetti, A., \& Mengoni, A. (2014, May). \textit{StreamingTrim 1.0: a Java software for dynamic trimming of 16S rRNA sequence data from metagenetic studies}. Paper presented at PhD Day\textsuperscript{5}. Sesto Fiorentino, Italy.

\end{itemize}

%The source files of this thesis will be available after the discussion as a git repository on \href{https://github.com/mgalardini/PhDGala}{GitHub} (@mgalardini).
